---
title: "Case Study of countries GDP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Presented by Michael Toolin, Venkat Karsala and Jamie Villenueva
#Introduction
A series of questions were presented regarding GDP data.  The data was provided in two different data sets.
The data sets needed to be merged and then the questions are answered

```{r missmatch, echo=TRUE}
dfgdp_2012_final <- read.csv('../data/dfgdp_2012_final.csv', stringsAsFactors = FALSE)
edstats_final <- read.csv('../data/edstats_final.csv', stringsAsFactors = FALSE)
NumMatches <- inner_join(dfgdp_2012_final,edstats_final)%>% summarise(cnt=n())%>% as.numeric()
``` 
The first question presented was how many ID's in the two data sets matched.  
```{r}
cat("There were",NumMatches,"found to be the same in the data sets \n")
```

```{r arranged, echo=TRUE}
cleaned_merged1 <- read.csv('../data/cleaned_merged_final.csv')
arrangByUSD <- arrange(cleaned_merged1, Millions.US.Dollars)
cat("The 13th country in the ascending GDP rankings is",arrangByUSD$Country[13]%>%as.character(),"\n")
```

The second question presented was to find the average GDP rankings for the **High Income:OECD** group and the **High Income:nonOECD** group.
<!--- insert R code to display the averages here --->

Below is plot of for **GDP of all countries** by Income Group:
<!--- insert R code to show the plot            --->
<!---            ```{r pressure, echo=TRUE}    --->
<!---               plot(pressure)              --->

The table below displays the GDP rankings seperated into 5 different groups.
<!--- insert R code to show the table of the 5 groups 
                  {r cars}
                  summary(cars)               --->

<!--- insert R code; how many countries in Lower middle group
      among the 38 nations with highest GDP           --->



