---
title: "Case Study of countries GDP"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## Presented by Michael Toolin, Venkat Karsala and Jamie Villenueva
#Introduction
A series of questions were presented regarding GDP data.  The data was provided in two different data sets.
The data sets needed to be merged and then the questions are answered

```{r missmatch, echo=TRUE}
dfgdp_2012_final <- read.csv('../data/dfgdp_2012_final.csv', stringsAsFactors = FALSE)
edstats_final <- read.csv('../data/edstats_final.csv', stringsAsFactors = FALSE)
NumMatches <- inner_join(dfgdp_2012_final,edstats_final)%>% summarise(cnt=n())%>% as.numeric()
``` 
The first question presented was how many ID's in the two data sets matched.  
```{r}
cat("There were",NumMatches,"found to be the same in the data sets \n")
```
The second question is find out the 13th Country when order by GDP in ascending order
```{r arranged, echo=TRUE}
cleaned_merged1 <- read.csv('../data/cleaned_merged_final.csv')
arrangByUSD <- arrange(cleaned_merged1, Millions.US.Dollars)
cat("The 13th country in the ascending GDP rankings is",arrangByUSD$Country[13]%>%as.character(),"\n")
```

The third question presented was to find the average GDP rankings for the **High Income:OECD** group and the **High Income:nonOECD** group.
```{r MeanRaning, echo=TRUE}
cleaned_merged1 %>% filter(Income.Group %in% c("High income: OECD" , "High income: nonOECD")) %>% group_by(Income.Group) %>% summarise( average = mean(Ranking)) %>% mutate_if(is.factor, as.character) %>% as.data.frame(as.matrix(),stringsAsFactors=F)
```

Below is plot of for **GDP of all countries** by Income Group:
```{r GDPPlot, echo=TRUE}
colgrp <- cleaned_merged1$Income.Group
ggplot(cleaned_merged1, aes(x=CountryCode, y=log(Millions.US.Dollars), group=colgrp,shape=colgrp)) + geom_point(aes(colour=colgrp)) + scale_shape_manual(values=c(19,20,21,22,23))+scale_colour_manual(values=c("blue", "red","gray","black","purple"))

ggplot(cleaned_merged1, aes(x=CountryCode, y=log(Millions.US.Dollars), group=colgrp,shape=colgrp)) + 
geom_bar(stat="identity",aes(colour=colgrp)) +
scale_shape_manual(values=c(19,20,21,22,23))+
scale_colour_manual(values=c("blue", "red","gray","black","purple"))
```

The table below displays the GDP rankings seperated into 5 different groups.
<!--- insert R code to show the table of the 5 groups 
                  {r cars}
                  summary(cars)               --->

<!--- insert R code; how many countries in Lower middle group
      among the 38 nations with highest GDP           --->



